### 2024-03-29 doc/sections/introduction.tex
Part of pushing through syntactic stuff

MAV and BV are \emph{deep inference} systems. Deep Inference~\cite{Guglielmi14:di} is generalisation of Gentzen's methodology for designing proof systems, which arose from Guglielmi's attempts to relate process algebra~\cite[CCS]{Milner80:CCS,Milner89:CC} to linear logic~\cite{Girard87:ll}.
The problem with such a relation is that, while the multiplicative connectives of linear logic capture parallel composition, no connective of linear logic captures \emph{sequential composition}.
Eventually, Guglielmi's attempts yielded BV, which extends multiplicative linear logic~\cite[MLL]{Girard87:ll} with a self-dual non-commutative connective that captures sequential composition.
Such a connective was already present in another extension of linear logic, Pomset logic~\cite{Retore97:pomset}, where it arose from the study of coherence space semantics of linear logic~\cite[\C8]{GirardTL89:proofs}.
Recently, Nguyễn and Stra{\ss}burger~\cite{NguyenS22:bvisnotpl} showed that, while BV is similar to Pomset logic, the two are not the same, as the theorems of BV form a proper subset of the theorems of Pomset logic.
Neither BV nor Pomset logic has a sequent calculus. Tiu~\cite{Tiu06:sisii} showed that sequent calculus cannot capture BV, and it is assumed that this result extends to Pomset logic.

>>>>

MAV and BV are \emph{deep inference} systems. Deep Inference~\cite{Guglielmi14:di} is generalisation of Gentzen's methodology for designing proof systems, which arose from Guglielmi's attempts to relate process algebra~\cite[CCS]{Milner80:CCS,Milner89:CC} to classical linear logic~\cite[CLL]{Girard87:ll}.
The problem with such a relation is that, while the multiplicative connectives of linear logic capture parallel composition, no connective of linear logic captures \emph{sequential composition}.
Eventually, Guglielmi's attempts yielded BV, which extends multiplicative linear logic~\cite[MLL]{Girard87:ll} with a self-dual non-commutative connective that captures sequential composition.
Such a connective was already present in another extension of linear logic, Pomset logic~\cite{Retore97:pomset}, where it arose from the study of coherence space semantics of linear logic~\cite[\C8]{GirardTL89:proofs}.
Recently, Nguyễn and Stra{\ss}burger~\cite{NguyenS22:bvisnotpl} showed that, while BV is similar to Pomset logic, the two are not the same, as the theorems of BV form a proper subset of the theorems of Pomset logic.
Neither BV nor Pomset logic has a sequent calculus. Tiu~\cite{Tiu06:sisii} showed that sequent calculus cannot capture BV, and it is assumed that this result extends to Pomset logic.

### 2024-03-29 doc/sections/mav-cut-elimination.tex
[ tex ] typos and small fixes
Part of pushing through syntactic stuff

Let $\Chu(\ClosedLowerSet{\NMAV}, \ClosedDay{I})$ be the MAV-algebra
constructed (Theorem \ref{thm:algebra-from-frame}) from the normal
proof MAV-frame (Proposition \ref{prop:nmav-frame}), where elements
are positive/negative pairs of $+$-closed lower sets of formulas. We
define the valuation of atoms as
$V(\alpha) = \ChuEmbed(\ClosedLowerEmbed(\vDual{\alpha}))$. By Theorem
\ref{thm:soundness}, we have an interpretation of MAV formulas
$\sem{P}$ such that if $P \longrightarrow^* I$, then
$I \sqsubseteq \sem{P}$. We now prove our main proposition about this
interpretation in the MAV-algebra derived from the MAV-frame of normal
proofs \NMAV{} that will allow us to derive the admissibility of all
the non-normal proof rules of MAV.

\begin{proposition}\label{prop:embedding-sem}
  $\sem{P} \sqsubseteq \lnot (\ChuEmbed(\ClosedLowerEmbed(P)))$
\end{proposition}

\begin{proof}
  By Definition \ref{defn:chu}, this statement is two inclusions
  between pairs of $+$-closed lower sets:
  \begin{enumerate}
  \item $\ClosedLowerEmbed(P) \subseteq \sem{\vP}^-$
  \item $\sem{\vP}^+ \subseteq \ClosedLowerEmbed(P) \rightblackspoon^+ \ClosedDay{I}$
  \end{enumerate}
  We prove the second assuming the first. It suffices to prove that
  $\sem{\vP}^+ \ClosedDay{\bullet} \ClosedLowerEmbed(P) \subseteq
  \ClosedDay{I}$, which follows from the first part and the property
  of all Chu-elements that
  $\sem{\vP}^+ \ClosedDay{\bullet} \sem{\vP}^- \subseteq \ClosedDay{I}$.

  We prove the first part by induction on $P$. In the cases when
  $P = \vUnit$ or $P = \vDual{\alpha}$, we already have
  $\sem{P}^- = \ClosedLowerEmbed(P)$. When $P = \alpha$, we have
  $\sem{\alpha}^- = \ClosedLowerEmbed(\vDual{\alpha})
  \rightblackspoon^+ \ClosedDay{I}$. It suffices to prove that
  $\ClosedLowerEmbed(\alpha) \ClosedDay{\bullet}
  \ClosedLowerEmbed(\vDual{\alpha}) \subseteq \ClosedDay{I}$, which follows
  from the preservation of monoid operations by $\ClosedLowerEmbed$
  and the \RuleLabel{Axiom} rule of MAV.

  When $P = Q \vParr R$, $Q \vWith R$, or $Q \vSeq R$, the result
  follows from preservation of the corresponding monoid structure by
  $\ClosedLowerEmbed$. For example,
  $\ClosedLowerEmbed(Q \vParr R) \subseteq \ClosedLowerEmbed(Q)
  \ClosedDay{\bullet} \ClosedLowerEmbed(R) \subseteq \sem{Q}^-
  \ClosedDay{\bullet} \sem{R}^- = \sem{Q \vParr R}^-$.

  When $P = Q \vPlus R$, we have
  $\ClosedLowerEmbed(Q \vPlus R) \subseteq \ClosedLowerEmbed(Q)$ and
  $\ClosedLowerEmbed(Q \vPlus R) \subseteq \ClosedLowerEmbed(R)$, by the
  \RuleLabel{Left} and \RuleLabel{Right} rules. Therefore,
  $\ClosedLowerEmbed(Q \vPlus R) \subseteq \ClosedLowerEmbed(Q) \lor
  \ClosedLowerEmbed(R) \subseteq \sem{Q}^- \lor \sem{R}^- = \sem{Q \vPlus
    R}^-$.

  When $P = Q \vTens R$, we have
  $\sem{Q \vTens R}^- = (\sem{Q}^+ \rightblackspoon \sem{R}^-) \land
  (\sem{R}^+ \rightblackspoon \sem{Q}^-)$. We prove inclusion in the
  left-hand side, the right-hand side is similar. The key property we
  need to prove is:
  \begin{equation}\label{eq:interaction-leq}
    \ClosedLowerEmbed(Q \vTens R) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
  \end{equation}
  Using the monoidality and monotonicity of $\alpha$, this inclusion
  is implied by the following inclusion in $\LowerSet{\NMAV}$:
  \begin{displaymath}
    \LowerEmbed(Q \vTens R) \Day{\bullet} (U(\ClosedLowerEmbed(Q)) \rightblackspoon \Day{I})
    \subseteq
    \LowerEmbed(R)
  \end{displaymath}
  which follows from the \RuleLabel{Switch} rule of MAV and
  calculation. Using \ref{eq:interaction-leq}, and inclusion (ii)
  above, we can prove the inequality we need:
  \begin{displaymath}
    \ClosedLowerEmbed(Q \vTens R) \ClosedDay{\bullet} \sem{Q}^+
    \subseteq
    \ClosedLowerEmbed(Q \vTens R) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
    \subseteq
    \sem{R}^-
  \end{displaymath}
  Using the residuation property of $\ClosedDay{\bullet}$ we can conclude.
\end{proof}

\begin{theorem}\label{thm:cut-elim}
  If $P \longrightarrow^* I$ in MAV, then there is a normal proof $P \longrightarrow^* I$.
\end{theorem}

\begin{proof}
  By Theorem \ref{thm:soundness}, $P \longrightarrow^* I$ in MAV
  implies $I \sqsubseteq \sem{P}$. Combined with Proposition
  \ref{prop:embedding-sem}, we have
  $I \sqsubseteq \lnot \ChuEmbed(\LowerEmbed(P)$. By Definition
  \ref{defn:chu} of the ordering of Chu elements, we have
  $\ClosedLowerEmbed(P) \subseteq \ClosedDay{I}$. Since
  $P \in \ClosedLowerEmbed(P)$, we have $P \in \ClosedDay{I}$, which
  by definition means that there is a normal proof
  $P \longrightarrow^* I$.
\end{proof}

>>>>

Let $\Chu(\ClosedLowerSet{\NMAV}, \ClosedDay{I})$ be the MAV-algebra
constructed (\Cref{thm:algebra-from-frame}) from the normal
proof MAV-frame (\Cref{prop:nmav-frame}), where elements
are positive/negative pairs of $+$-closed lower sets of formulas. We
define the valuation of atoms as
$V(\va) = \ChuEmbed(\ClosedLowerEmbed(\vDual\va))$. By
\Cref{thm:soundness}, we have an interpretation of MAV formulas
$\sem{\vP}$ such that if $\vP\vInferFrom*\vUnit$, then
$I \sqsubseteq \sem{\vP}$. We now prove our main proposition about this
interpretation in the MAV-algebra derived from the MAV-frame of normal
proofs \NMAV{} that will allow us to derive the admissibility of all
the non-normal proof rules of MAV.

\begin{proposition}\label[prop]{prop:embedding-sem}
  $\sem{\vP} \sqsubseteq \lnot (\ChuEmbed(\ClosedLowerEmbed(P)))$
\end{proposition}

\begin{proof}
  By \Cref{defn:chu}, this statement is two inclusions
  between pairs of $+$-closed lower sets:
  \begin{enumerate}
    \item $\ClosedLowerEmbed(P) \subseteq \sem{\vP}^-$
    \item $\sem{\vP}^+ \subseteq \ClosedLowerEmbed(P) \rightblackspoon^+ \ClosedDay{I}$
  \end{enumerate}
  We prove the second assuming the first. It suffices to prove that
  $\sem{\vP}^+ \ClosedDay{\bullet} \ClosedLowerEmbed(P) \subseteq
    \ClosedDay{I}$, which follows from the first part and the property
  of all Chu-elements that
  $\sem{\vP}^+ \ClosedDay{\bullet} \sem{\vP}^- \subseteq \ClosedDay{I}$.

  We prove the first part by induction on $\vP$.
  In the cases when $\vP = \vUnit$ or $\vP = \vDual\va$, we already have
  $\sem{\vP}^- = \ClosedLowerEmbed(P)$. When $\vP = \va$, we have
  $\sem{\va}^- = \ClosedLowerEmbed(\vDual\va)
    \rightblackspoon^+ \ClosedDay{I}$. It suffices to prove that
  $\ClosedLowerEmbed(\va) \ClosedDay{\bullet}
    \ClosedLowerEmbed(\vDual\va) \subseteq \ClosedDay{I}$, which
  follows from the preservation of monoid operations by
  $\ClosedLowerEmbed$ and the \cref{rule:AtomAxiom} rule.

  When $\vP = \vQ\vParr\vR$, $\vQ\vWith\vR$, or $\vQ\vSeq\vR$, the result
  follows from preservation of the corresponding monoid structure by
  $\ClosedLowerEmbed$. For example,
  $\ClosedLowerEmbed(\vQ\vParr\vR) \subseteq \ClosedLowerEmbed(\vQ)
    \ClosedDay{\bullet} \ClosedLowerEmbed(\vR) \subseteq \sem{\vQ}^-
    \ClosedDay{\bullet} \sem{\vR}^- = \sem{\vQ\vParr\vR}^-$.

  When $\vP = \vQ\vPlus\vR$, we have
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vQ)$ and
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vR)$, by the
  \cref{rule:Left} and \cref{rule:Right} rules. Therefore,
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vQ) \lor
    \ClosedLowerEmbed(\vR) \subseteq \sem{\vQ}^- \lor \sem{\vR}^- = \sem{\vQ\vPlus\vR}^-$.

  When $\vP = \vQ\vTens\vR$, we have
  $\sem{\vQ\vTens\vR}^- = (\sem{\vQ}^+ \rightblackspoon \sem{\vR}^-) \land
    (\sem{\vR}^+ \rightblackspoon \sem{\vQ}^-)$. We prove inclusion in the
  left-hand side, the right-hand side is similar. The key property we
  need to prove is:
  \begin{equation}\label{eq:interaction-leq}
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
  \end{equation}
  Using the monoidality and monotonicity of $\va$, this inclusion
  is implied by the following inclusion in $\LowerSet{\NMAV}$:
  \begin{displaymath}
    \LowerEmbed(\vQ\vTens\vR) \Day{\bullet} (U(\ClosedLowerEmbed(Q)) \rightblackspoon \Day{I})
    \subseteq
    \LowerEmbed(R)
  \end{displaymath}
  which follows from the \cref{rule:Switch} rule of MAV and
  calculation. Using \ref{eq:interaction-leq}, and inclusion (ii)
  above, we can prove the inequality we need:
  \begin{displaymath}
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} \sem{\vQ}^+
    \subseteq
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
    \subseteq
    \sem{\vR}^-
  \end{displaymath}
  Using the residuation property of $\ClosedDay{\bullet}$ we can conclude.
\end{proof}

\begin{theorem}\label[thm]{thm:cut-elim}
  If $\vP\vInferFrom*\vUnit$ in MAV, then there is a normal proof $\vP\vInferFrom*\vUnit$.
\end{theorem}

\begin{proof}
  By \Cref{thm:soundness}, $\vP\vInferFrom*\vUnit$ in MAV
  implies $I \sqsubseteq \sem{\vP}$. Combined with \Cref{prop:embedding-sem}, we have
  $I \sqsubseteq \lnot \ChuEmbed(\LowerEmbed(\vP)$. By
  \Cref{defn:chu} of the ordering of Chu elements, we have
  $\ClosedLowerEmbed(\vP) \subseteq \ClosedDay{I}$. Since
  $P \in \ClosedLowerEmbed(\vP)$, we have $\vP \in \ClosedDay{I}$, which
  by definition means that there is a normal proof
  $\vP\vInferFrom*\vUnit$.
\end{proof}

Another consequence of \Cref{prop:embedding-sem} is that
MAV is complete for the MAV-frame semantics:

\begin{theorem}\label[thm]{thm:completeness}
  MAV is complete for the MAV-frame semantics: if, for all MAV-frames
  $F$, $\sem{\vUnit} \sqsubseteq \sem{\vP}$ in
  $(\Chu(\ClosedLowerSet{F}, \ClosedDay{I}), \sqsubseteq)$, then
  $\vP\vInferFrom*\vUnit$.
\end{theorem}

### 2024-03-29 doc/sections/mav-mechanisation.tex
Part of pushing through syntactic stuff

As well as giving confidence in our results, a key benefit of
formalising the development in a proof assistant like Agda is that the
proof normalisation process (Theorem \ref{thm:cut-elim}) is
executable. For example, we can normalise the one step SMAV proof
(using \RuleLabel{Axiom}) of $\vP \vParr \vDual{\vP}$ where
$\vP = (\vUnit \vPlus \vUnit) \vSeq (\vUnit \vWith \vUnit)$ to a FIXME
step MAV proof, given in the Appendix.

>>>>

As well as giving confidence in our results, a key benefit of
formalising the development in a proof assistant like Agda is that the
proof normalisation process (\Cref{thm:cut-elim}) is
executable. For example, we can normalise the one step SMAV proof
(using \cref{rule:Axiom}) of $\vP \vParr \vDual{\vP}$ where
$\vP = (\vUnit \vPlus \vUnit) \vSeq (\vUnit \vWith \vUnit)$ to a FIXME
step MAV proof, given in the Appendix.
